---
title: "Preliminary NDVI analyses, Denver area"
author: "Michael D Garber"
output: bookdown::html_document2 #the rest will get picked up from the _site.yml
---

This document presents preliminary analyses characterizing the NDVI of certain places and census tracts throughout the Denver area. 


```{r loadpackages, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
#First, load packages used in this Rmarkdown doc.
library(here)
library(tidyverse)
library(sf)
library(terra)
library(raster) 
library(mapview)
library(knitr) #to make sure kable works
library(viridis)
library(scales)
library(lubridate) #for ggplot
```

# NDVI for the Denver area, 2016-2021
We obtained the normalized difference vegetation index (NDVI) from the Landsat-8 satellite at a resolution of 30 meters squared at a roughly 15-day interval between 2016 and 2021 for the following area around Denver:

## Define study area
```{r denverbbox, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
load("den_metro_bbox_custom.RData")
#around Denver and Jefferson Counties but only as far south as Castle Rock, i.e, not all of Jeff Co to limit data gathered
mv_den_metro_bbox_custom = den_metro_bbox_custom %>% mapview(layer.name = "Study area")
mv_den_metro_bbox_custom
```

We gathered this data using the [rgee package](https://cran.r-project.org/web/packages/rgee/index.html), an R package that facilitates connection to the Google Earth Engine via Python. Please see these two scripts for specific details on the process for gathering and processing the NDVI data in this area:

https://github.com/michaeldgarber/green-space-denver/blob/main/scripts/1a_get_landsat_ndvi_denver.R

https://github.com/michaeldgarber/green-space-denver/blob/main/scripts/1b_wrangle_check_landsat_ndvi_denver.R

## NDVI for one day
Here is the NDVI over the study area on a cloud-free day, July 4, 2021:
```{r terragoodmo, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}
#Update 5/26/22 this may not work with mapview anymore, so use plot instead for speed and
setwd(here("data-processed"))
pal_terrain_col = rev(terrain.colors(100)) 
ndvi_den_metro_terr_5_yr = terra::rast("ndvi_den_metro_terr_5_yr.tif") 

ndvi_den_metro_terr_5_yr$`20210704_NDVI` %>% plot()
#old code:
# mv_good_date =ndvi_den_metro_terr_5_yr$`20210704_NDVI` %>% 
#   raster::raster() %>%  
#   mapview(
#     col.regions = pal_terrain_col,
#     layer.name = "NDVI, July 4, 2021")
# mv_good_date

```


# Characterize NDVI of places of interest
## First, determine valid (cloud-free) dates
### Testing a few places with expected high NDVI
First, pick a few places where we would expect NDVI to be high in the summer on a cloud-free day. If it's not high, then we can assume the image is bad (i.e., clouds in the way). These test places were determined by examining the cloud-free day (July 4, 2021) above. We chose three areas: an area east of Evergreen, Colorado; a plot in City Park; and a plot in the Indian Tree Golf Club:

(These areas are defined here: https://github.com/michaeldgarber/green-space-denver/blob/main/scripts/1b_wrangle_check_landsat_ndvi_denver.R)

```{r loadtestareas, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
getwd()
load("bbox_evergreen_east.RData")
load("bbox_indian_tree_golf.RData")
load("bbox_city_park.RData")
load("den_metro_bbox_custom.RData")
mv_evergreen_east = bbox_evergreen_east %>%
  mapview(col.regions = "red",  layer.name = "Evergreen East")
mv_indian_tree_golf = bbox_indian_tree_golf %>%
  mapview(col.regions = "red", layer.name = "Indian Tree Golf")
mv_city_park =bbox_city_park %>%
  mapview(col.regions = "red",  layer.name = "City Park plot")
mv_den_metro_bbox_custom = den_metro_bbox_custom %>% 
  mapview(layer.name = "Study area", alpha.regions = .2)
mv_den_metro_bbox_custom+ mv_evergreen_east + mv_indian_tree_golf + mv_city_park
```

### What is NDVI over time for these test places?
Let's look at the temporal NDVI trend for these three places, without excluding any dates. This is the mean NDVI on that day. That is, each place includes several pixels of size 30 meters squared. This is the average of the NDVI for those pixels for that place on that day.
```{r testplacesggplot, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
load("ndvi_test_places_day_wrangle.RData")
ndvi_test_places_day_wrangle %>% 
  ggplot(
    aes(
      x=date, 
      y=ndvi_mean  
    ))+
  geom_line(aes(colour = test_place_name))+
  geom_point(aes(colour = test_place_name))+
  scale_x_date(labels=date_format("%Y-%m-%d"), date_breaks = "3 months")+
  theme(axis.text.x = element_text(angle = 45, hjust=1))
```

Hmm, if we look closely, we can see that there are some very low NDVI values for some of these places even in the summer, which is not plausible. Those low values must indicate obstruction by clouds. For example, look at a day when NDVI was measured as very low in City Park despite it being mid-May when we would expect it to be higher. So I'm going to exclude dates like these. 
```{r mapviewbaddate, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
pal_viridis_trunc=viridis::viridis_pal(end=.9) #trunc for truncated
ndvi_den_metro_terr_5_yr = terra::rast("ndvi_den_metro_terr_5_yr.tif")
ndvi_den_metro_terr_5_yr$`20180517_NDVI` %>% #Use 2018-05-17
  plot()

#old code that uses mapview. decide to use plot instead. see abobve
# mv_bad_date =ndvi_den_metro_terr_5_yr$`20180517_NDVI` %>% #Use 2018-05-17
#   raster::raster() %>%  
#   mapview(
#     col.regions = pal_viridis_trunc,
#     layer.name = "NDVI, May 17, 2018")
# mv_bad_date
```

The specific NDVI thresholds for each thresholds are noted in this script (https://github.com/michaeldgarber/green-space-denver/blob/main/scripts/1b_wrangle_check_landsat_ndvi_denver.R) and warrant further discussion. 

The valid dates are restricted to May, June, July, and August and specifically are:
```{r printvaliddates, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
load("lookup_date_is_valid_all.RData") #load dataset of valid dates created elsewhere
lookup_date_is_valid_all %>% 
  filter(date_is_valid_all==1) %>% 
  dplyr::select(date) %>% 
  pull()
```


## Load places of interest

Load the places-of-interest dataset, which is created here: https://github.com/michaeldgarber/green-space-denver/blob/main/scripts/0_read_denver_native_zones.R

```{r loaddata, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
load("places_native_geo.RData")
load("places_native_nogeo.RData")
load("native_places_ndvi_day_nogeo.RData")
```

Print information about the places of interest, including approximate percent nativity where available.
```{r printtibble, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE, results = 'asis'}
places_native_nogeo %>% 
  dplyr::select(place_name, native_percent, place_type, place_id) %>% 
  kable()
```


### Map places of interest by type
Map these places by type (nativity spectrum or high-plant diversity).
```{r mapviewplacesnative, eval=T, echo=FALSE, warning=FALSE, message=FALSE}
#I need a version that is just points because the polygons are hard to see on the map
#comment 6/2/22 this is causing a recursion error:
#https://stackoverflow.com/questions/14719349/error-c-stack-usage-is-too-close-to-the-limit
#Just don't evaluate it.
library(here)
library(tidyverse)
library(sf)
library(mapview)
setwd(here("data-processed"))
load("places_native_geo.RData")
load("places_native_nogeo.RData")
load("native_places_ndvi_day_nogeo.RData")
sf::sf_use_s2(FALSE) 
places_native_geo_points = places_native_geo %>% 
  st_make_valid() %>% 
  st_transform(2876) %>% 
  st_centroid()

mv_places_native_polygons = places_native_geo %>%
  st_simplify() %>% 
  st_transform(2876) %>% 
  st_union(by_feature = TRUE) %>% 
  st_cast("POLYGON") %>% #This fixed the c-stack issue. too many combined sf feature types.
  st_as_sf() %>% 
  mapview(zcol = "place_type",
          layer.name = "Place category (as polygons)")


mv_places_native_points = places_native_geo_points %>%
  mapview(zcol = "place_type",
          layer.name = "Place category (as points)")

library(leaflet)
library(leaflet.extras)
mv_places_polygons_points =mv_places_native_points+
  mv_places_native_polygons

mv_places_polygons_points@map %>%
  addFullscreenControl()

#This works, but I don't need to run every time.
# setwd(here("results"))
# getwd()
#mapshot(mv_places_polygons_points, file = "mv_places_polygons_points.png")

```

## Graph NDVI over time for these places
During valid dates only (above).

### For the places with percent nativity values
The first set of polygons sent over.
```{r ggplotndvinative, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}
native_places_ndvi_day_nogeo %>% 
  filter(date_is_valid_all==1) %>% 
  filter(place_type == "native spectrum") %>% 
  ggplot(aes(
    x=date, 
    y=ndvi_med # median
  ))+
  geom_ribbon( #ribbon around 25th and 75th percentile
    aes(ymin =ndvi_25, ymax = ndvi_75 ),
    alpha=.4
  )+
  ylab("NDVI, Median") +
  xlab("Date") +
  geom_line( size=.7 ) +#note size better than lwd
  geom_point()+
  scale_x_date(breaks = pretty_breaks())+
  scale_y_continuous(
    limits = c(0, NA),  #force axis origin to be zero
    breaks= seq(0,0.8,by = 0.1))+ 
  theme(
    axis.text.x=element_text(angle=60, hjust=1),
        panel.border = element_rect(
          colour = "gray72", size = 0.5, fill=NA))+
  facet_grid(   # facet them
    cols = vars(place_name_fac),
    labeller = label_wrap_gen() #wrap facet labels
  )
```


### For the places with high plant diversity
This is the second set of polygons sent.

```{r ggplotndvidivers, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}
native_places_ndvi_day_nogeo %>% 
  filter(date_is_valid_all==1) %>% 
  filter(place_type == "high diversity") %>% 
  ggplot(aes(
    x=date, 
    y=ndvi_med # median
  ))+
  geom_ribbon( 
    aes(ymin =ndvi_25, ymax = ndvi_75 ),
    alpha=.4
  )+
  ylab("NDVI, Median") +
  xlab("Date") +
  geom_line( size=.7 ) + 
  geom_point()+
  scale_x_date(breaks = pretty_breaks())+
  scale_y_continuous(
    limits = c(0, NA),  
    breaks= seq(0,0.8,by = 0.1))+ 
  theme(
    axis.text.x=element_text(angle=60, hjust=1),
        panel.border = element_rect(
          colour = "gray72", size = 0.5, fill=NA))+
  facet_grid(   
    cols = vars(place_name_fac),
    labeller = label_wrap_gen() #wrap facet labels
  )
```


## Graph NDVI against percent nativity
Among the first set of polygons with values for percent nativity.
```{r ggplotscatter, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}
native_places_ndvi_day_nogeo %>% 
  filter(date_is_valid_all==1) %>% 
  filter(place_type == "native spectrum") %>% 
  ggplot(aes(
    x=native_percent, 
    y=ndvi_med # median
  ))+
  geom_point(
    aes(colour=place_name_fac), size = 1.5, 
    alpha=.5 #varying alpha to illustrate density
    ) +
  xlab("Percent native") +
  ylab("NDVI, median") +
  scale_y_continuous(
    limits = c(0, NA),  #force axis origin to be zero
    breaks= seq(0,0.8,by = 0.1))+ 
  scale_color_hue(
    name = "Place name"
  )
```


## Summary of NDVI for these places
```{r summariseoverall, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE, results = 'asis'}
setwd(here("data-processed"))
load("places_native_nogeo.RData")
load("native_places_ndvi_day_nogeo.RData")
native_places_ndvi_day_nogeo %>% 
  filter(date_is_valid_all==1) %>% 
  group_by(place_type, place_name_fac) %>% 
  summarise(
    ndvi_mean = mean(ndvi_mean_wt, na.rm=TRUE), # mean of means
    ndvi_25th = quantile(ndvi_med, probs =c(0.25), na.rm=TRUE), #percentile of medians
    ndvi_med = median(ndvi_med, na.rm=TRUE), #percentile of medians
    ndvi_75th = quantile(ndvi_med, probs =c(0.75), na.rm=TRUE) ) %>% #percentile of medians
  ungroup() %>% 
  left_join(places_native_nogeo, by = c("place_name_fac", "place_type")) %>% 
  dplyr::select(
    starts_with("place_type"),
    starts_with("place_name_fac"),
    starts_with("area_ft2"),
    starts_with("area_mi"),
    starts_with("ndvi_mea"),
    starts_with("ndvi_25"),
    starts_with("ndvi_med"),
    starts_with("ndvi_75th"),
  ) %>% 
  as_tibble() %>% 
  knitr::kable(
    booktabs = TRUE,
    col.names = c("Place type", "Place name", "Area (sq. ft.)", "Area (sq. mi.)", "NDVI, mean", "NDVI, 25th-ile", "NDVI, 50th-ile", "NDVI, 75th-ile"),
  digits = c(0, 0, 0, 3, 2, 2, 2, 2, 2)
  )
```


# NDVI of census tracts
In this section, we summarize NDVI for census tracts in Denver County. Before characterizing NDVI of each census tract, we removed bodies of water. NDVI values of water approach -1, and we did not want to "penalize" a census tract for having large amounts of water.


Here is a map of the median NDVI of the census tracts in the area on 2021-07-04:
```{r , eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
#4/8/22 note I used to have another code chunk above this one showing a different day (2021-06-10), but the code was breaking.
library(here)
setwd(here("data-processed"))
load("den_co_tract_ndvi_day_geo.RData")
pal_terrain_col = rev(terrain.colors(100)) 
#I have a version with "tracts" saved. If I re-run the code, "tract" should work.
den_co_tract_ndvi_day_geo %>% 
  filter(date == "2021-07-04") %>% 
  mapview(
    layer.name = "NDVI, Median",
     col.regions = pal_terrain_col,
    zcol = "ndvi_med" 
    )
```

Here is a histogram of the distribution of median NDVI for these census tracts on June 10, 2021:
```{r ndvitractshisto2021-06-10, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}
den_co_tract_ndvi_day_geo %>% 
  filter(date == "2021-06-10") %>% 
  ggplot(aes(ndvi_med))+
  geom_histogram()
```


# NDVI of public green spaces
In this section, we examine the NDVI levels of public green space in this area, including city parks and nature reserves. Please see this code for more details on how we gathered the polygons for the greens spaces: https://github.com/michaeldgarber/green-space-denver/blob/main/scripts/0_load_denver_osm_parks_water.R

Here are the parks we downloaded:
```{r parkbytype, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
load("den_jeff_co_green_space_public.RData")
den_jeff_co_green_space_public %>% 
  mapview(zcol = "osm_value", layer.name = "Green space type")
```

Like with census tracts, we removed bodies of water from these green spaces before measuring NDVI. 
Here is the NDVI on June 10, 2021:
```{r ndviparks2021-06-10, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
load("den_metro_green_space_ndvi_day_geo.RData")
load("den_jeff_co_geo.RData")
pal_viridis_trunc=viridis::viridis_pal(end=.9) #trunc for truncated
mv_den_jeff_co_geo = den_jeff_co_geo %>%  
  mapview(
    layer.name = "County name",
    color = c("red", "orange"),
    col.regions = c("red", "orange"),
    lwd=2,
    zcol = "county_name_short", alpha.regions = 0)
mv_den_metro_green_space_ndvi_day_geo = den_metro_green_space_ndvi_day_geo %>%
  dplyr::select( #limit some of the variables before mapview
    starts_with("osm"),
    starts_with("ndvi"),
    starts_with("date"),
    starts_with("county")) %>% 
  filter(date == "2021-06-10") %>% 
  mapview(
    layer.name = "NDVI, Median",
    col.regions = pal_viridis_trunc,
    zcol = "ndvi_med" 
  )
mv_den_metro_green_space_ndvi_day_geo+mv_den_jeff_co_geo
```

Most parks are in the city, and most protected areas are in the mountains. Summarize NDVI by type of green space:
```{r summarise_by_park_type, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE, results = 'asis'}
setwd(here("data-processed"))
library(knitr)
load("den_metro_green_space_ndvi_day_nogeo.RData")
den_metro_green_space_ndvi_day_nogeo %>% 
  filter(date_is_valid_all==1) %>% 
  group_by(osm_value) %>% 
  summarise(
    ndvi_25th = quantile(ndvi_med, probs =c(0.25), na.rm=TRUE),
    ndvi_med = median(ndvi_med, na.rm=TRUE),
    ndvi_75th = quantile(ndvi_med, probs =c(0.75), na.rm=TRUE),
    ) %>% 
  ungroup() %>% 
  knitr::kable()
```