---
title: "Population density and NDVI: possible measures considering both"
author: "Michael D Garber"
date: '2022-05-24'
#always_allow_html: true
#note word doesn't seem to render unless I render the whole site ()rmarkdown::render_site(encoding = 'UTF-8')
#update 7/8/22 I removed the word output
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_depth: 4
    include: 
      after_body: footer.html
#  bookdown::word_document2:
#    toc: true 
---

This page presents some ideas for expressing NDVI in terms of population density for the City of Denver, Colorado.

```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, include = FALSE}
library(tidyverse)
library(sf)
library(mapview)
library(here)
library(terra)
library(raster)
library(leaflet)
library(leaflet.extras)
setwd(here("data-processed"))
load("den_co_bg_pop_per_ndvi.RData") #made in scripts/4_pop_dens_per_ndvi.R
suppressMessages(source(here("scripts", "0_read_equity_indices.R"))) #easier than loading the requisite data 
```

# Population density
A map of population density by block group per 2016-2020 American Community Survey 5-year Estimates
```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
mv_den_co_bg_pop_per_ndvi =den_co_bg_pop_per_ndvi %>% 
  mapview(zcol = "pop_dens_mi2_all",
    layer.name = "Popuplation density per square mile")

mv_den_co_bg_pop_per_ndvi@map %>% 
  addFullscreenControl()
```
Population density tertiles
```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
pal_terrain_3_cat = rev(terrain.colors(4)[1:3])
den_co_bg_pop_per_ndvi %>% 
  mapview(
    layer.name = "Population density tertile",
    zcol = "pop_dens_mi2_tertile")
```

# NDVI
A map of NDVI of each block group. Note NDVI is measured using Landsat-8 at a spatial resolution of 30 meters squared on July 4, 2021, a cloud-free day. The weighted average NDVI of each block group excludes bodies of water.
```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
pal_terrain = terrain.colors(100) %>% rev()#reverse the order of the palette
mv_den_co_bg_ndvi  = den_co_bg_pop_per_ndvi %>% 
  dplyr::select(bg_fips, ndvi_mean_wt) %>% 
  mapview(
    layer.name = "NDVI, block group",
    zcol = "ndvi_mean_wt",
    col.regions = pal_terrain, 
    at = seq(-0.1, 1, 0.1)
  )
mv_den_co_bg_ndvi@map %>% 
  addFullscreenControl()
```

NDVI tertiles, overall
```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
den_co_bg_pop_per_ndvi %>% 
  mapview(
    layer.name = "NDVI tertile",
    col.regions = pal_terrain_3_cat, 
    zcol = "ndvi_tertile_overall")
```

# Explore relationship between population density, NDVI, and equity indices
## Scatterplot of population density x NDVI
```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
den_co_bg_pop_per_ndvi %>% 
  ggplot(aes(x=pop_dens_mi2_all, y=ndvi_mean_wt )) + 
  geom_point(aes(colour=pop_dens_mi2_tertile ), size = 1) +
  ylab("NDVI") +
  xlab("Pop. density") +
  theme_bw(base_size = 14) +
  scale_color_hue(
    name = "Pop. density tertile"
  )
```


## Scatterplot of equity index (CDPHE definition) x NDVI x pop. density tertile
```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
den_co_bg_pop_per_ndvi %>% 
  dplyr::select(-tract_fips) %>% 
  link_equity_indices() %>% 
  ggplot(aes(x=equity_bg_cdphe, y=ndvi_mean_wt )) + 
  geom_point(aes(colour=pop_dens_mi2_tertile ), size = 1) +
  ylab("NDVI") +
  xlab("CDPHE Equity Index (higher is worse)") +
  theme_bw(base_size = 14) +
  scale_color_hue(
    name = "Pop. density tertile"
  )
```

## Scatterplot of equity index (Denver definition) x NDVI
```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
den_co_bg_pop_per_ndvi %>% 
  dplyr::select(-tract_fips) %>% 
  link_equity_indices() %>% 
  ggplot(aes(x=equity_nbhd_denver, y=ndvi_mean_wt )) + 
  geom_point(aes(colour=pop_dens_mi2_tertile ), size = 1) +
  ylab("NDVI") +
  xlab("Denver Equity Index (lower is worse)") +
  theme_bw(base_size = 14) +
  scale_color_hue(
    name = "Pop. density tertile"
  )
```

## Scatterplot of equity index (CDPHE definition) x pop. density
```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
den_co_bg_pop_per_ndvi %>% 
  dplyr::select(-tract_fips) %>% 
  link_equity_indices() %>% 
  ggplot(aes(x=equity_bg_cdphe, y=pop_dens_mi2_all )) + 
  geom_point(size = 1) +
  ylab("Pop. density") +
  xlab("CDPHE Equity Index (higher is worse)") +
  theme_bw(base_size = 14) 
```

# Possible composite measures considering both population density and NDVI
1. Population density per NDVI (population density per square mile / NDVI)
2. A ranking of population density per NDVI
3. Tertiles of NDVI within tertiles of population density

## Population density per NDVI
The first is difficult to visualize on a map because the distribution has a right skew with some very high values and most much lower.
```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
summary(den_co_bg_pop_per_ndvi$pop_dens_per_ndvi)
den_co_bg_pop_per_ndvi %>% 
  ggplot(aes(pop_dens_per_ndvi))+
  geom_histogram()+
  theme_bw(base_size = 14) +
  xlab("Pop. density per NDVI")
```

## Ranking of population density per NDVI
The second is easier to visualize because the distribution is uniform by definition (a simple rank).
```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
summary(den_co_bg_pop_per_ndvi$pop_dens_per_ndvi_rank)
den_co_bg_pop_per_ndvi %>% 
  ggplot(aes(pop_dens_per_ndvi_rank))+
  geom_histogram()+
  theme_bw(base_size = 14) +
  xlab("Block group ranking of pop. density per NDVI")
```

Map population density per NDVI and the ranking of population density per NDVI on the same map.
```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
library(viridis)
library(scales)
mv_pop_dens_per_ndvi=den_co_bg_pop_per_ndvi %>% 
  mapview(
    col.regions = viridis_pal(option = "C"),
    layer.name = "Pop. density per NDVI",
    zcol = "pop_dens_per_ndvi")

mv_rank=den_co_bg_pop_per_ndvi %>% 
  mapview(
    col.regions = viridis_pal(option = "C"),
    layer.name = "Rank of pop. density per NDVI",
    zcol = "pop_dens_per_ndvi_rank")

mv_both = mv_pop_dens_per_ndvi +
  mv_rank

mv_both@map %>% 
  leaflet.extras::addFullscreenControl()
```


## NDVI tertiles within pop. density tertiles
### Map of NDVI tertiles within tertiles of population density
```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
#Create the filters of each pop density tertile
pop_dens_mi2_tertile_1 = den_co_bg_pop_per_ndvi %>% 
  st_set_geometry(NULL) %>% 
  group_by(pop_dens_mi2_tertile) %>% 
  summarise(n=n()) %>% 
  dplyr::select(pop_dens_mi2_tertile) %>% 
  ungroup() %>% 
  dplyr::slice(1) %>% 
  pull()

pop_dens_mi2_tertile_2 = den_co_bg_pop_per_ndvi %>% 
  st_set_geometry(NULL) %>% 
  group_by(pop_dens_mi2_tertile) %>% 
  summarise(n=n()) %>% 
  dplyr::select(pop_dens_mi2_tertile) %>% 
  ungroup() %>% 
  dplyr::slice(2) %>% 
  pull()

pop_dens_mi2_tertile_3 = den_co_bg_pop_per_ndvi %>% 
  st_set_geometry(NULL) %>% 
  group_by(pop_dens_mi2_tertile) %>% 
  summarise(n=n()) %>% 
  dplyr::select(pop_dens_mi2_tertile) %>% 
  ungroup() %>% 
  dplyr::slice(3) %>% 
  pull()

#See mapview issue regarding a bug in getting the labels to print.
#That's why I'm simply calling the categorie 1, 2, 3, which is less informative
#than including the actual range of values, but this will work for now.
pal_terrain_3_cat = rev(terrain.colors(4)[1:3])
mv_ndvi_tertile_pop_dens_1 = den_co_bg_pop_per_ndvi %>% 
  filter(pop_dens_mi2_tertile== pop_dens_mi2_tertile_1) %>% 
  mapview(
    layer.name = "NDVI tertile within pop. dens. tertile 1 (lowest)",
    col.regions = pal_terrain_3_cat, 
    zcol = "ndvi_tertile_within_pop_dens_tertile_label")

mv_ndvi_tertile_pop_dens_2 =den_co_bg_pop_per_ndvi %>% 
  filter(pop_dens_mi2_tertile== pop_dens_mi2_tertile_2) %>% 
  mapview(
    layer.name = "NDVI tertile within pop. dens tertile 2 (middle)",
    col.regions = pal_terrain_3_cat, 
    zcol = "ndvi_tertile_within_pop_dens_tertile_label")

mv_ndvi_tertile_pop_dens_3 =den_co_bg_pop_per_ndvi %>% 
  filter(pop_dens_mi2_tertile== pop_dens_mi2_tertile_3) %>% 
  mapview(
    layer.name = "NDVI tertile within pop. dens. tertile 3 (highest)",
    col.regions = pal_terrain_3_cat, 
    zcol = "ndvi_tertile_within_pop_dens_tertile_label")

mv_ndvi_tertile_overall =den_co_bg_pop_per_ndvi %>% 
  mapview(
    layer.name = "NDVI tertile, overall",
    col.regions = pal_terrain_3_cat, 
    zcol = "ndvi_tertile_overall")

mv_ndvi_pop_dens_tertile_combined = mv_ndvi_tertile_pop_dens_1 + 
  mv_ndvi_tertile_pop_dens_2+
  mv_ndvi_tertile_pop_dens_3 + 
  mv_ndvi_tertile_overall

mv_ndvi_pop_dens_tertile_combined@map %>% 
  leaflet.extras::addFullscreenControl()
```

### Maps of ratios and differences of relative NDVI within tertiles of population density (highest NDVI tertile is referent)

Ratio of block group NDVI compared with the median NDVI of the highest NDVI tertile in that population density tertile
```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
library(viridis)
library(RColorBrewer)
library(shades)
setwd(here("data-processed"))
load("pop_density_tertile_den_co_bg.RData")
load("den_co_bg_pop_per_ndvi.RData")
greys = RColorBrewer::brewer.pal(n=3, "Greys")
mv_pop_density_tertile = pop_density_tertile_den_co_bg %>% 
  mutate(
    pop_dens_mi2_med_round = round(pop_dens_mi2_med)
  ) %>% 
  mapview(
    layer.name = "Median pop. density within pop. dens. tertile",
    col.regions = greys,
  zcol = "pop_dens_mi2_med_round")

mv_ratio_ndvi_med_pop_dens_ndvi_tert=  den_co_bg_pop_per_ndvi %>% 
  mapview(
    col.regions = viridis_pal(option = "plasma", direction=-1),
    layer.name = "Ratio of NDVI with median of highest NDVI tertile",
    zcol = "ratio_ndvi_med_pop_dens_ndvi_tert")
mv_ratio_both =mv_ratio_ndvi_med_pop_dens_ndvi_tert+mv_pop_density_tertile
mv_ratio_both@map %>% 
  leaflet.extras::addFullscreenControl()
```


Difference of block group NDVI compared with the median NDVI of the highest NDVI tertile in that population density tertile
```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
library(viridis)
library(RColorBrewer)
library(shades)
mv_pop_density_tertile = pop_density_tertile_den_co_bg %>% 
  mutate(
    pop_dens_mi2_med_round = round(pop_dens_mi2_med)
  ) %>% 
  mapview(
    layer.name = "Median pop. density within pop. dens. tertile",
    col.regions = greys,
  zcol = "pop_dens_mi2_med_round")
mv_diff_ndvi_med_pop_dens_ndvi_tert=  den_co_bg_pop_per_ndvi %>% 
  mapview(
    col.regions = viridis_pal(option = "plasma", direction=-1),
    layer.name = "Difference of NDVI with median of highest NDVI tertile",
    zcol = "diff_ndvi_med_pop_dens_ndvi_tert")

mv_diff_both = mv_diff_ndvi_med_pop_dens_ndvi_tert + mv_pop_density_tertile
mv_diff_both@map %>% 
  leaflet.extras::addFullscreenControl()
```
