---
title: "NATURGREEN paper tables & figure (main text)"
author: "Michael D Garber"
date: '2022-05-24'
#always_allow_html: true
#note word doesn't seem to render unless I render the whole site ()rmarkdown::render_site(encoding = 'UTF-8')
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_depth: 4
    include: 
      after_body: footer.html
  bookdown::word_document2:
    toc: true 
---

This document contains tables and figures presented in the main text. The URL for this page is https://michaeldgarber.github.io/green-space-denver/tables_figs_main_text.html

Additional detail is presented in the appendix:
https://michaeldgarber.github.io/green-space-denver/tables_figs_appendix.html

7/8/22 Update to concatenate (paste0) confidence intervals, so they're all in the same cell.

```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
library(here)
library(tidyverse)
library(mapview)
library(sf)
library(leaflet.extras)
```

# Figure 1
Note this is directly copied from the appendix version.
The weighted block-group-level mean NDVI on July 4, 2021 is presented here.

```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.cap = "Mean normalized difference vegetation index (NDVI) of census block groups in Denver, Colorado (excluding some in the northeast, described in main text) on July 4, 2021 as measured by the Landsat-8 satellite on July 4, 2021. NDVI is measured at a spatial resolution of 30 meters squared. The weighted block-group-level mean NDVI is weighted by the proportion Landsat pixels overlap the water-free land area of the block group"}
#update may 11 2022 to add figure caption
setwd(here("data-processed"))
load("den_co_bg_ndvi_geo.RData") #updated april 18 2022
#names(den_co_bg_ndvi_geo)
pal_terrain = terrain.colors(100) %>% rev()#reverse the order of the palette
mv_den_co_bg_ndvi  = den_co_bg_ndvi_geo %>% 
  dplyr::select(bg_fips, ndvi_mean_wt) %>% 
  mapview(
    layer.name = "NDVI, block group",
    zcol = "ndvi_mean_wt",
    col.regions = pal_terrain, 
    at = seq(-0.1, 1, 0.1)
  )
mv_den_co_bg_ndvi@map %>% 
  addFullscreenControl()
```

# Tables 
## Table 1: Description of scenarios (area, pop affected)
```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
load("hia_all_over_ndvi_over_equity_est_boot.RData")
load("hia_all_over_ndvi_by_equity_est_boot.RData")
load("lookup_scenario_sort_order.RData") #to get it to sort how I want
load("lookup_scenario_main_text.RData") #to restrict to those in the main text if necessary
load("lookup_scenario_num.RData")
load("lookup_scenario_label.RData")
library(scales)
hia_all_over_ndvi_over_equity_est_boot %>% 
  bind_rows(hia_all_over_ndvi_by_equity_est_boot) %>% 
  arrange(scenario_sort_order) %>% 
  #filter to just the main sub-scenarios to present in the main text
  left_join(lookup_scenario_main_text, by = c("scenario", "scenario_sub")) %>% 
  filter(scenario_main_text==1) %>% 
  left_join(lookup_scenario_num , by = c("scenario", "scenario_sub")) %>% #add this for presentation
  left_join(lookup_scenario_label, by=  c("scenario", "scenario_sub")) %>% #add better scenario labels
  #Create concatenated version of estimates + confidence intervals in one cell
  #I learned about formatC, which I have never used before 
  #https://stackoverflow.com/questions/3443687/formatting-decimal-places-in-r
  mutate(
    area_mi2_bg_int_tx_conc = paste0(
      as.character(formatC(area_mi2_bg_int_tx_med,digits=1, format = "f")),
      " (", 
      as.character(formatC(area_mi2_bg_int_tx_min,digits=1, format = "f")),
      ", ",
      as.character(formatC(area_mi2_bg_int_tx_max,digits=1, format="f")),
      ")")
    ,
    area_mi2_bg_int_res_conc = paste0(
      as.character(formatC(area_mi2_bg_int_res_med,digits=1, format = "f")),
      " (", 
      as.character(formatC(area_mi2_bg_int_res_min,digits=1, format = "f")),
      ", ",
      as.character(formatC(area_mi2_bg_int_res_max,digits=1, format="f")),
      ")")
    ,
    area_prop_tx_res_conc = paste0(
      as.character(formatC(area_prop_tx_res_med,digits=2, format = "f")),
      " (", 
      as.character(formatC(area_prop_tx_res_min,digits=2, format = "f")),
      ", ",
      as.character(formatC(area_prop_tx_res_max,digits=2, format="f")),
      ")"),
    #here, use big.mark to add a comma.
    pop_affected_conc =  paste0(
        as.character(formatC(pop_affected_med,digits=0, format = "f", big.mark=",")),
        " (", 
        as.character(formatC(pop_affected_ll,digits=0, format = "f", big.mark=",")),
        "; ", #using a semi-colon as a separator instead to avoid confusion with comma
        #for every thousand.
        as.character(formatC(pop_affected_ul,digits=0, format="f", big.mark=",")),
        ")")
  ) %>% 
  dplyr::select(
    scenario_num,
    scenario_label, 
  #  scenario_sub, #not presenting sub-scenarios in the main text table
    starts_with("equity_nbhd_denver_tertile"),
    ends_with("_conc") #this picks up all the variables you just created with formatC
  ) %>%
  ungroup() %>% 
  knitr::kable(
    caption = "The cumulative area of the treatment, its corresponding residential buffer, and the population affected",
    booktabs = TRUE,
    format.args = list(big.mark = ","),
#    digits = c(0, 0, 0, 1, 1, 1, 1, 1, 1,2,2,2,0, 0,0),
    col.names = c(
            "Scenario num.",
      "Scenario",
      "Equity Tertile",
      "Area, treatment (mi^2^), median (95% UI)",
      "Area, residential buffer (mi^2^), median (95% UI)",
      "Prop. of residential buffer treated, median (95% UI)",
      "Pop. affected, estimate (95% UI)"
      )  )
```
## Table 2: NDVI at baseline, alternative, and difference therein
```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
load("hia_all_over_ndvi_over_equity_est_boot.RData")
load("hia_all_over_ndvi_by_equity_est_boot.RData")
load("lookup_scenario_sort_order.RData") #to get it to sort how I want
load("lookup_scenario_main_text.RData") #to restrict to those in the main text if necessary
load("lookup_scenario_num.RData")
hia_all_over_ndvi_over_equity_est_boot %>% 
  bind_rows(hia_all_over_ndvi_by_equity_est_boot) %>% 
  arrange(scenario_sort_order) %>% 
  #filter to just the main sub-scenarios to present in the main text
  left_join(lookup_scenario_main_text, by = c("scenario", "scenario_sub")) %>% 
  filter(scenario_main_text==1) %>% 
  left_join(lookup_scenario_num , by = c("scenario", "scenario_sub")) %>% #add this for presentation
  left_join(lookup_scenario_label, by=  c("scenario", "scenario_sub")) %>% #add better scenario labels
    #more concatenate
  mutate(
    
    #baseline ndvi of the treatment area 
    ndvi_mean_wt_tx_conc = paste0(
      as.character(formatC(ndvi_mean_wt_tx_med,digits=2, format = "f")),
      " (", 
      as.character(formatC(ndvi_mean_wt_tx_min,digits=2, format = "f")),
      ", ",
      as.character(formatC(ndvi_mean_wt_tx_max,digits=2, format="f")),
      ")"),
    
    #baseline NDVI of the full residential area
    ndvi_quo_conc = paste0(
      as.character(formatC(ndvi_quo_med,digits=2, format = "f")),
      " (", 
      as.character(formatC(ndvi_quo_min,digits=2, format = "f")),
      ", ",
      as.character(formatC(ndvi_quo_max,digits=2, format="f")),
      ")")
    ,
    ndvi_mean_alt_conc = paste0(
      as.character(formatC(ndvi_mean_alt_med,digits=2, format = "f")),
      " (", 
      as.character(formatC(ndvi_mean_alt_min,digits=2, format = "f")),
      ", ",
      as.character(formatC(ndvi_mean_alt_max,digits=2, format="f")),
      ")"),
    ndvi_diff_conc = paste0(
      as.character(formatC(ndvi_diff_med,digits=3, format = "f")),
      " (", 
      as.character(formatC(ndvi_diff_min,digits=3, format = "f")),
      ", ",
      as.character(formatC(ndvi_diff_max,digits=3, format="f")),
      ")")
  ) %>% 
  dplyr::select(
    scenario_num,
    scenario_label, 
    starts_with("equity_nbhd_denver_tertile"),
    ends_with("_conc") #this picks up all the variables you just created with formatC
  ) %>% 
  knitr::kable(
    caption = "Baseline and alternative NDVI of the residential buffer and the corresponding difference",
    booktabs = TRUE,
#    format.args = list(big.mark = ","), #not necessary because we formatted in the paste0
#    digits = c(0,0,0,2,2,2,2,2,2,2,2,2,3,3,3), ##max digits after decimal. also not necessary
    col.names = c(
      "Scenario num.",
      "Scenario", 
      "Equity Tertile", 
      "Baseline NDVI, treatment area, median (95% UI)",
      "Baseline NDVI, residential buffer, median (95% UI)",
      "Alternative NDVI, residential buffer, median (95% UI)",
      "Difference: NDVI, residential buffer, median (95% UI)"
      )
) 
```

## Table 3: Estimated deaths averted (total and rate per 100k) under each scenario
```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
load("hia_all_over_ndvi_over_equity_est_boot.RData")
load("hia_all_over_ndvi_by_equity_est_boot.RData")
load("lookup_scenario_sort_order.RData") #to get it to sort how I want
load("lookup_scenario_main_text.RData") #to restrict to those in the main text if necessary
load("lookup_scenario_num.RData")
hia_all_over_ndvi_over_equity_est_boot %>% 
  bind_rows(hia_all_over_ndvi_by_equity_est_boot) %>% 
  arrange(scenario_sort_order) %>% 
  #filter to just the main sub-scenarios to present in the main text
  left_join(lookup_scenario_main_text, by = c("scenario", "scenario_sub")) %>% 
  filter(scenario_main_text==1) %>% 
  left_join(lookup_scenario_num , by = c("scenario", "scenario_sub")) %>% #add this for presentation
  left_join(lookup_scenario_label, by=  c("scenario", "scenario_sub")) %>% #add better scenario labels
    #more concatenate
  mutate(
    deaths_prevented_count_conc = paste0(
      as.character(formatC(deaths_prevented_count_med,digits=0, format = "f")),
      " (", 
      as.character(formatC(deaths_prevented_count_ll,digits=0, format = "f")),
      ", ",
      as.character(formatC(deaths_prevented_count_ul,digits=0, format="f")),
      ")"),
    
    deaths_prevented_per_pop_100k_conc = paste0(
      as.character(formatC(deaths_prevented_per_pop_100k_med,digits=0, format = "f")),
      " (", 
      as.character(formatC(deaths_prevented_per_pop_100k_ll,digits=0, format = "f")),
      ", ",
      as.character(formatC(deaths_prevented_per_pop_100k_ul,digits=0, format="f")),
      ")")
  ) %>% 
  dplyr::select(
    scenario_num,
    scenario_label, 
    starts_with("equity_nbhd_denver_tertile"),
    ends_with("conc")
  ) %>% 
  knitr::kable(
    caption = "Estimated annual premature deaths prevented",
    booktabs = TRUE,
        col.names = c(
      "Scenario num.",
      "Scenario", 
      "Equity Tertile", 
      "Deaths prevented, estimate (95% UI)", 
      "Deaths prevented per 100k, estimate (95% UI)"
      )
  )
```


