---
title: "NATURGREEN paper tables & figure (appendix)"
author: "Michael D Garber"
date: '2022-05-24'
#always_allow_html: true
#note word doesn't seem to render unless I render the whole site ()rmarkdown::render_site(encoding = 'UTF-8').
output:
  bookdown::word_document2:
    toc: true 
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_depth: 4
    include: 
      after_body: footer.html
bibliography: references.bib
csl: american-medical-association.csl
link-citations: true
---

Filename: tables_figs_appendix

This document contains supplementary material, mostly tables and figures, as well as some additional text and description. This web page could plausibly be rendered as a PDF or Word Doc for submission.

The tables and figures intended for the main text are here:

https://michaeldgarber.github.io/green-space-denver/tables_figs_main_text.html

As of now, this also contains all of the information presented in the main text. That is, the tables and figures presented in the main text are a subset of those presented here.

# Denver demographics

## Table of total population, population density, age, race, and poverty of study area

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
library(here)
library(tidyverse)
library(mapview)
library(sf)
#all of this info is created here:
#scripts/0_import_manage_denver_acs.R
setwd(here("data-processed"))
load("lookup_tract_nbhd_northeast_exclude.RData")
load("den_tract_acs5_wrangle_geo.RData")
#names(den_tract_acs5_wrangle_geo)
#use tract-level data for the overall table
den_tract_acs5_wrangle_geo %>% 
  filter(county_fips=="031") %>% 
  left_join(lookup_tract_nbhd_northeast_exclude, by = "tract_fips") %>% 
#  filter(nbhd_northeast_exclude==0) %>% #Decide not to exclude, actually
  st_set_geometry(NULL) %>% 
  group_by(county_fips) %>% 
  summarise(
    area_mi2 = sum(area_mi2_tract, na.rm=TRUE),
    pop_tot = sum(pop_tot, na.rm=TRUE),
    
    age_med = median(age_med , na.rm=TRUE),
    hh_inc_med = median(hh_inc_med, na.rm=TRUE),
    
    race_tot = sum(race_tot, na.rm=TRUE),
    race_w = sum(race_w, na.rm=TRUE),
    race_h = sum(race_h, na.rm=TRUE),
    race_b = sum(race_b, na.rm=TRUE),
    
    pov_last_12 = sum(pov_last_12, na.rm=TRUE),
    pov_last_12_tot = sum(pov_last_12_tot, na.rm=TRUE),

  ) %>% 
  ungroup() %>% 
  mutate(
    pop_dens_mi2 = pop_tot/area_mi2,
    race_w_prop = race_w/race_tot,           #prop for proportion
    race_b_prop = race_b/race_tot,
    race_h_prop = race_h/race_tot, 
    race_nw_prop = 1-race_w_prop, #nw for nonwhite (i.e, 1-white)
    race_o_prop = 1-(race_b_prop + race_w_prop +race_h_prop),
    
    pov_last_12_prop = pov_last_12/pov_last_12_tot  #percent poverty
  ) %>% 
  dplyr::select(
    area_mi2, pop_tot, pop_dens_mi2, race_w_prop, race_b_prop, race_h_prop, race_o_prop,
    age_med,
    pov_last_12_prop,
    hh_inc_med
  ) %>% 
  pivot_longer(
    cols = everything(),
    names_to = "Measure"
    ) %>% 
  knitr::kable(
    booktabs = TRUE,
    caption = "Summary of study-area demographics",
    digits = 2)

```

## Map population density by block group

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
#scripts/0_import_manage_denver_acs.R
setwd(here("data-processed"))
load("lookup_tract_nbhd_northeast_exclude.RData")
load("den_bg_acs5_wrangle_geo.RData")
den_bg_acs5_wrangle_geo %>% 
  filter(county_fips=="031") %>% 
  left_join(lookup_tract_nbhd_northeast_exclude, by = "tract_fips") %>% 
#  filter(nbhd_northeast_exclude==0) %>% 
  dplyr::select(ends_with( "fips"), contains("pop_dens")) %>% 
  mapview(
    layer.name = "Population density per square mile",
    zcol = "pop_dens_mi2_all")
```

## Map median age by census tract

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
load("lookup_tract_nbhd_northeast_exclude.RData")
load("den_bg_acs5_wrangle_geo.RData")
load("den_tract_acs5_wrangle_geo.RData")
den_tract_acs5_wrangle_geo %>% 
  filter(county_fips=="031") %>% 
  left_join(lookup_tract_nbhd_northeast_exclude, by = "tract_fips") %>% 
 # filter(nbhd_northeast_exclude==0) %>% 
  dplyr::select(ends_with( "fips"), contains("age")) %>% 
  mapview(
    layer.name = "Median age",
    zcol = "age_med")
```

## Map race (proportion non-white) by census tract

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
load("lookup_tract_nbhd_northeast_exclude.RData")
load("den_tract_acs5_wrangle_geo.RData")
den_tract_acs5_wrangle_geo %>% 
  filter(county_fips=="031") %>% 
  left_join(lookup_tract_nbhd_northeast_exclude, by = "tract_fips") %>% 
#  filter(nbhd_northeast_exclude==0) %>% 
  dplyr::select(ends_with( "fips"), contains("race")) %>% 
  mapview(
    layer.name = "Proportion non-white race",
    zcol = "race_nw_prop")
```

## Map percent in poverty in last 12 months by census tract

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
load("lookup_tract_nbhd_northeast_exclude.RData")
load("den_tract_acs5_wrangle_geo.RData")
den_tract_acs5_wrangle_geo %>% 
  filter(county_fips=="031") %>% 
  left_join(lookup_tract_nbhd_northeast_exclude, by = "tract_fips") %>% 
#  filter(nbhd_northeast_exclude==0) %>% 
  dplyr::select(ends_with( "fips"), contains("pov")) %>% 
  mapview(
    layer.name = "Proportion in poverty, last year",
    zcol = "pov_last_12_prop")
```

## Map median household income by census tract

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
load("lookup_tract_nbhd_northeast_exclude.RData")
load("den_tract_acs5_wrangle_geo.RData")
den_tract_acs5_wrangle_geo %>% 
  filter(county_fips=="031") %>% 
  left_join(lookup_tract_nbhd_northeast_exclude, by = "tract_fips") %>% 
 # filter(nbhd_northeast_exclude==0) %>% 
  dplyr::select(ends_with( "fips"), contains("hh_inc")) %>% 
  mapview(
    layer.name = "Median household income",
    zcol = "hh_inc_med")
```

# Native-plants polygons

## Map of native-plants polygons

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
#I need a version that is just points because the polygons are hard to see on the map
library(here)
library(tidyverse)
library(sf)
library(terra)
library(raster) 
library(mapview) 
library(knitr) #to make sure kable works
library(viridis)
library(scales)
library(lubridate) #for ggplot
setwd(here("data-processed"))
load("places_native_geo.RData")
load("places_native_nogeo.RData")
load("native_places_ndvi_day_nogeo.RData")
sf::sf_use_s2(FALSE) 
places_native_geo_points = places_native_geo %>% 
  st_make_valid() %>% 
  st_transform(2876) %>% 
  st_centroid()

mv_places_native_polygons = places_native_geo %>%
  st_simplify() %>% 
  st_transform(2876) %>% 
  st_union(by_feature = TRUE) %>% 
  st_cast("POLYGON") %>% #This fixed the c-stack issue. too many combined sf feature types.
  st_as_sf() %>% 
  mapview(zcol = "place_type",
          layer.name = "Place category (as polygons)")


mv_places_native_points = places_native_geo_points %>%
  mapview(zcol = "place_type",
          layer.name = "Place category (as points)")

library(leaflet)
library(leaflet.extras)
mv_places_polygons_points =mv_places_native_points+
  mv_places_native_polygons

mv_places_polygons_points@map %>%
  addFullscreenControl()

```

\newpage

## NDVI of native-plants polygons

```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
library(here)
library(tidyverse)
library(rmarkdown)
library(sf)
setwd(here("data-processed"))
load("places_native_geo.RData")
load("places_native_nogeo.RData")
load("native_places_ndvi_day_nogeo.RData")
native_places_ndvi_day_nogeo %>% 
  filter(date_is_valid_all==1) %>% 
  group_by(place_type, place_name_fac) %>% 
  summarise(
    ndvi_mean = mean(ndvi_mean_wt, na.rm=TRUE), # mean of means
    ndvi_25th = quantile(ndvi_med, probs =c(0.25), na.rm=TRUE), #percentile of medians
    ndvi_med = median(ndvi_med, na.rm=TRUE), #percentile of medians
    ndvi_75th = quantile(ndvi_med, probs =c(0.75), na.rm=TRUE) ) %>% #percentile of medians
  ungroup() %>% 
  left_join(places_native_nogeo, by = c("place_name_fac", "place_type")) %>% 
  dplyr::select(
    starts_with("place_type"),
    starts_with("place_name_fac"),
    starts_with("area_ac"),
    starts_with("area_mi"),
    starts_with("ndvi_mea"),
    starts_with("ndvi_25"),
    starts_with("ndvi_med"),
    starts_with("ndvi_75th"),
  ) %>% 
  as_tibble() %>% 
  knitr::kable(
    booktabs = TRUE,
        caption = "NDVI of selected places of varying levels of plant nativity on selected cloud-free summer (May, June, July, August) days over five years (2016-2021)",
    col.names = c("Place type", "Place name", "Area (acres)", "Area (sq. mi.)", "NDVI, mean", "NDVI, 25th-ile", "NDVI, 50th-ile", "NDVI, 75th-ile"),
  digits = c(0, 0, 1, 2, 2, 2, 2, 2, 2)
  )
```

# Equity indices

Both the State of Colorado and the City of Denver have created equity indices to guide their work. The Colorado Department of Public Health & Environment (CDPHE) describes their equity index here[@cdphe2021] and makes the index available for download here.[@cdphe] The City of Denver Department of Public Health & Environment (DDPHE) describes the neighborhood-level equity index here[@ddphe] and makes the data available for download here.[@denver]

Note that for the CDPHE index, a higher score is a worse score, whereas for the DDPHE index, a lower score is a worse score. The map below presents tertiles of each score.

```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.cap = "Comparing equity indices"}
library(scales)
setwd(here("data-processed"))
load("den_equity_ind_2020.RData") #updated april 18 2022
load("cdphe_equity_geo.RData")
mv_cdphe_equity_den_tertile =  cdphe_equity_geo %>% 
  filter(county_fips == "031") %>% 
  mapview(
    layer.name = "Equity Index, CDPHE, tertile",
    zcol = "equity_bg_cdphe_tertile_den",
    col.regions = viridis_pal(direction = -1) #flip scale
    )

mv_den_equity_ind_tertile = den_equity_ind_2020 %>% 
  mapview(
    layer.name = "Equity Index, Denver, tertile",
    zcol = "equity_nbhd_denver_tertile",
    col.regions = viridis_pal(direction = 1) #don't flip scale
    
    )
mv_equity_both = mv_den_equity_ind_tertile + mv_cdphe_equity_den_tertile
library(leaflet.extras)
mv_equity_both@map %>% 
  addFullscreenControl()
```

Below are results of a chi-square test comparing tertiles of CDPHE Equity Index vs the DDPHE Equity index.

```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
#to keep consistent, using exact same function used in 3_hia_for_each_scenario
link_equity_indices= function(df){
  df %>% 
    left_join(lookup_colorado_bg_tract, by = "bg_fips") %>% #2020 ok
    left_join(lookup_den_nbhd_tract, by = "tract_fips") %>%  
    left_join(lookup_equity_nbhd_denver, by = "nbhd_id") %>% 
    rename(bg_fips_2020 = bg_fips) %>%   
    left_join(lookup_colorado_bg_2020_vs_2019_nogeo, by = "bg_fips_2020") %>% 
    left_join(lookup_equity_bg_cdphe_2019, by = "bg_fips_2019") %>% 
    rename(bg_fips = bg_fips_2020)      #change name back
}

setwd(here("data-processed"))
load("den_bg_acs5_wrangle_geo.RData")
load("lookup_colorado_bg_tract.RData")
load("lookup_den_nbhd_tract.RData") #neighborhood ids in denver linked to tracts
load("lookup_equity_nbhd_denver.RData")
load("lookup_colorado_bg_2020_vs_2019_nogeo.RData")
load("lookup_equity_bg_cdphe_2019.RData")
compare_tertiles = den_bg_acs5_wrangle_geo %>% 
  dplyr::select(bg_fips, county_fips, geometry) %>% 
  link_equity_indices() %>% 
  filter(county_fips == "031") %>% 
  dplyr::select(contains("fips"), contains("equity")) %>% 
  st_set_geometry(NULL)

compare_tertiles_chisq = chisq.test(table(
  compare_tertiles$equity_bg_cdphe_tertile_den, compare_tertiles$equity_nbhd_denver_tertile)
)

compare_tertiles_chisq
```

# Visualization of scenarios

Note to self: This section is mostly copied from \~green-space-denver/docs/hia-results.Rmd

## Scenario 1: fixed proportion of all block groups

The weighted block-group-level mean NDVI on July 4, 2021 is presented here.

```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.cap = "Mean normalized difference vegetation index (NDVI) of census block groups in Denver, Colorado (excluding some in the northeast, described in main text) on July 4, 2021 as measured by the Landsat-8 satellite on July 4, 2021. NDVI is measured at a spatial resolution of 30 meters squared. The weighted block-group-level mean NDVI is weighted by the proportion Landsat pixels overlap the water-free land area of the block group"}
#update may 11 2022 to add figure caption
setwd(here("data-processed"))
load("den_co_bg_ndvi_geo.RData") #updated april 18 2022
#names(den_co_bg_ndvi_geo)
pal_terrain = terrain.colors(100) %>% rev()#reverse the order of the palette
mv_den_co_bg_ndvi  = den_co_bg_ndvi_geo %>% 
  dplyr::select(bg_fips, ndvi_mean_wt) %>% 
  mapview(
    layer.name = "NDVI, block group",
    zcol = "ndvi_mean_wt",
    col.regions = pal_terrain, 
    at = seq(-0.1, 1, 0.1)
  )
mv_den_co_bg_ndvi@map %>% 
  addFullscreenControl()
```

For each definition of native-plants NDVI, We excluded census blocks whose baseline NDVI was above the threshold. This map visualizes the NDVI cutoffs both for lower of the two values, that of the Denver Botanic Gardens Green Roof, and the higher of the two, that of Denver Botanic Gardens Native Plot.

```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
library(RColorBrewer)
setwd(here("data-processed"))
load("den_co_bg_ndvi_alt_all_nogeo.RData")
load("lookup_den_co_bg_no_wtr_geo.RData")
ndvi_below_thresh_pal = RColorBrewer::brewer.pal(3, "RdYlGn")[c(1,3)]  %>% rev()
mv_bg_below_native_threshold_low =den_co_bg_ndvi_alt_all_nogeo %>% 
  filter(ndvi_native_threshold<0.41) %>% #can pick either side of this.
  filter(scenario_sub == "20-pct") %>% 
  left_join(lookup_den_co_bg_no_wtr_geo, by = "bg_fips") %>% 
  st_as_sf() %>% 
  mapview(
    layer.name = "NDVI below lower native NDVI value",
    zcol = "ndvi_below_native_threshold_char",
    col.regions = ndvi_below_thresh_pal
  )

mv_bg_below_native_threshold_high =den_co_bg_ndvi_alt_all_nogeo %>% 
  filter(ndvi_native_threshold>0.45) %>% #can pick either side of this.
  filter(scenario_sub == "20-pct") %>% 
  left_join(lookup_den_co_bg_no_wtr_geo, by = "bg_fips") %>% 
  st_as_sf() %>% 
  mapview(
    layer.name = "NDVI below higher native NDVI value",
    zcol = "ndvi_below_native_threshold_char",
    col.regions = ndvi_below_thresh_pal
  )

mv_both_thresholds = mv_bg_below_native_threshold_high +
  mv_bg_below_native_threshold_low

mv_both_thresholds@map %>% 
  addFullscreenControl()
```

## Scenario 2: riparian areas

We measured NDVI in a 200-foot buffer, a 100-foot buffer, and a 50-foot buffer around all bodies of water in Denver. We downloaded bodies of water from OpenStreetMap (code [here](https://github.com/michaeldgarber/green-space-denver/blob/main/scripts/0_load_denver_osm_parks_water.R)). We defined residential exposure to these riparian areas as those individuals living within a 500-meter buffer.

The below map depicts mean NDVI in the portions of census block groups that intersect a 200-foot buffer as well as those pieces that intersect the part of the 500 m buffer that would not be intervened upon, i.e., the part between the 200-foot buffer and the edge of the 500 m residential buffer.

```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
load("den_bg_int_wtr_200ft_comp_ndvi.RData")
load("den_bg_int_wtr_200ft_tx_ndvi.RData")
load("den_co_osm_wtr.RData")
load("study_area_2876.RData")
pal_terrain_col = rev(terrain.colors(100)) 
mv_den_bg_int_wtr_200ft_comp_ndvi= den_bg_int_wtr_200ft_comp_ndvi %>% 
  mapview(
    col.regions = pal_terrain_col,
    at = seq(-.1, 1, 0.1), #0 to 1 by .1; define the breaks so consistent between layers
    layer.name = "NDVI, 200 ft - 500 m",
    zcol = "ndvi_mean_wt")

mv_den_bg_int_wtr_200ft_tx_ndvi= den_bg_int_wtr_200ft_tx_ndvi %>% 
  mapview(
    col.regions = pal_terrain_col,
    at = seq(-.1, 1, 0.1), #0 to 1 by .1; define the breaks so consistent between layers
    layer.name = "NDVI, 200 ft",
    zcol = "ndvi_mean_wt")

mv_den_co_osm_wtr = den_co_osm_wtr %>% 
  st_intersection(study_area_2876) %>% 
  mapview(
    layer.name = "Water type",
    col.regions = rainbow(n_distinct(den_co_osm_wtr$water_type)),
    zcol = "water_type")

#### Visualize all at once--------
mv_rip_all3 = mv_den_bg_int_wtr_200ft_tx_ndvi+ mv_den_bg_int_wtr_200ft_comp_ndvi +mv_den_co_osm_wtr
mv_rip_all3@map %>% 
  addFullscreenControl()
```

## Scenario 3: green infrastructure

### Scenario 3.1. stormwater retention projects

We were provided a list of planned projects throughout Denver. Per our conversations, we assume that about 75% of the project's footprint would consist of native or vegetation. Like for riparian areas, we defined exposure to the projects as those individuals living within a 500 m buffer of the projects. A map of baseline NDVI of the projects themselves and of a 500 m residential buffer around the projects is below.

```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
load("den_bg_int_ogi_proj_comp_ndvi.RData")
load("den_bg_int_ogi_proj_tx_ndvi.RData")
load("den_co_osm_wtr.RData")
load("study_area_2876.RData")
pal_terrain_col = rev(terrain.colors(100)) 

#### 500 mBuffer excluding the project---------
mv_den_bg_int_ogi_proj_comp_ndvi= den_bg_int_ogi_proj_comp_ndvi %>% 
  mapview(
    col.regions = pal_terrain_col,
    at = seq(-.1, 1, 0.1), #0 to 1 by .1; define the breaks so consistent between layers
    layer.name = "NDVI, 500m buffer, projects",
    zcol = "ndvi_mean_wt")

#### The project only----------
mv_den_bg_int_ogi_proj_tx_ndvi= den_bg_int_ogi_proj_tx_ndvi %>% 
  mapview(
    col.regions = pal_terrain_col,
    at = seq(-.1, 1, 0.1), #0 to 1 by .1; define the breaks so consistent between layers
    layer.name = "NDVI, projects only",
    zcol = "ndvi_mean_wt")

#### Bodies of water--------
load("den_co_osm_water.RData")
mv_den_co_osm_water = den_co_osm_water %>% 
  st_intersection(study_area_2876) %>% 
  mapview(
    layer.name = "Water type",
    col.regions = rainbow(n_distinct(den_co_osm_water$water_type)),
    zcol = "water_type")

#### Visualize all at once--------------
mv_ogi_proj_all3 = mv_den_bg_int_ogi_proj_comp_ndvi+ 
  mv_den_bg_int_ogi_proj_tx_ndvi +mv_den_co_osm_wtr
mv_ogi_proj_all3@map %>% 
  addFullscreenControl()
```

### Scenario 3.2. Site new or re-development storwmater controls

A regulatory scenario suggested by this agency is stormwater controls on new and re-development to offset the negative impacts to flooding and water quality due to the increase in impervious surfaces added by development. These stormwater control measures would add vegetation using native or adapted plants. The approach for estimating the health impact of the redevelopment controls is distinct in that we will not know where the re-developments will occur. We thus simulated possible locations. Per our conversations with the stakeholder, we anticipate the following number of parcels will be subject to these rules per year:

-   greater than 1 acre: about 100 sites per year
-   0.5-1.0 acre: about 25 sites per year
-   less than 0.5 acre: 400 sites per year

We gathered data on parcels from Denver's Open Data Portal ([Existing Landuse 2018](https://www.denvergov.org/opendata/dataset/city-and-county-of-denver-existing-landuse-2018)) and measured their area. A subset of these parcels near Union Station is mapped below. (Mapping all parcels is too much data to present.)

```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
library(here)
setwd(here("data-processed"))
load("den_landuse_union_sta.RData") #created scripts/0_read_wrangle_denver_land_use.R

mv_landuse_union_sta = den_landuse_union_sta %>% 
  dplyr::select(-starts_with("address")) %>% 
  mapview(
    zcol = "CPD_LANDUS",
    layer.name = "Land-use type",
    col.regions = rainbow(n=n_distinct(den_landuse_union_sta$CPD_LANDUS)))
mv_landuse_union_sta@map %>% 
  addFullscreenControl()
```

Then, from all parcels, we sampled 100 sites of size greater than 1 acre, 25 sites of size 0.5-1 acre, and 400 sites less than 0.5 acres. One such sample appears below.

```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
library(here)
setwd(here("data-processed"))
load("den_landuse_sample.RData")
mv_den_landuse_sample = den_landuse_sample %>% 
  mapview(
    layer.name = "Parcel size category",
    zcol = "parcel_size_cat")
mv_den_landuse_sample@map %>% 
  addFullscreenControl()
```

From this point, we followed the same framework as for other scenarios. The baseline NDVI of the sampled parcels and that of their corresponding 500m buffers is visualized below.

```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
load("den_bg_int_parcel_comp_ndvi.RData")
load("den_bg_int_parcel_tx_ndvi.RData")
load("study_area_2876.RData")
load("union_station_1_mi_4326.RData") # we will limit to a 1-mi radius around union station
pal_terrain_col = rev(terrain.colors(100)) 
library(leaflet.extras)
library(mapview)
#### Buffer excluding the parking---------
sf::sf_use_s2(FALSE) #have to use this as I was getting an error that loop not valid
mv_union_sta_den_bg_int_parcel_comp_ndvi= den_bg_int_parcel_comp_ndvi %>% 
  mapview(
    col.regions = pal_terrain_col,
    layer.name = "NDVI, 500 m buffer, parcel",
    zcol = "ndvi_mean_wt")

#### The parking only----------
mv_union_sta_den_bg_int_parcel_tx_ndvi= den_bg_int_parcel_tx_ndvi %>% 
  mapview(
    col.regions = pal_terrain_col,
    layer.name = "NDVI, parcel only",
    zcol = "ndvi_mean_wt")

#### Visualize all at once--------------
mv_parcel_both = mv_union_sta_den_bg_int_parcel_comp_ndvi+ 
  mv_union_sta_den_bg_int_parcel_tx_ndvi

mv_parcel_both@map %>% 
  addFullscreenControl()
```

## Scenario 4: parking lots

Finally, we propose replacing a portion of parking-lot surface with native plants. Denver's area is about 9% parking lot (below table). We obtained spatial data on parking lots from the City of Denver Open Data Catalog.[@denvero2016]

```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
load("den_prkng_tx_marg.RData")
den_prkng_tx_marg %>% 
  st_set_geometry(NULL) %>% 
  mutate(
    den_area_mi2_total = 155,
    prop_den_parking = area_mi2_prkng/den_area_mi2_total
         ) %>% 
  dplyr::select(-starts_with("area_ft2")) %>% 
  knitr::kable(
    booktabs = TRUE,
    caption = "Proportion of Denver's area covered by parking lots",
    col.names = c("Total area (mi^2^), parking lots", "Total area (mi^2^), Denver", "proportion parking lots"),
    digits = 2)
```

We similarly measured the NDVI on the parking lots as well as the NDVI in the census block groups within a 500-m buffer radius of any parking lot (which is almost the whole city). We present a subset of these areas in a 2-mile radius around Union Station (the complete data are too large to upload):

```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
load("den_bg_int_prkng_comp_ndvi.RData")
load("den_bg_int_prkng_tx_ndvi.RData")
load("study_area_2876.RData")
load("union_station_1_mi_4326.RData") # we will limit to a 1-mi radius around union station
load("union_station_1_mi_2876.RData")
load("union_station_2_mi_4326.RData") 
load("union_station_2_mi_2876.RData")

pal_terrain_col = rev(terrain.colors(100)) 
library(leaflet.extras)
library(mapview)
#### Buffer excluding the parking---------
sf::sf_use_s2(FALSE) #have to use this as I was getting an error that loop not valid
mv_union_sta_den_bg_int_prkng_comp_ndvi= den_bg_int_prkng_comp_ndvi %>% 
  st_transform(2876) %>% 
  st_simplify(dTolerance = 20) %>%  #make size much smaller for visual. intentionally ordered here
  st_intersection(union_station_2_mi_2876) %>% 
  st_make_valid() %>% 
  st_buffer(0) %>% 
  mapview(
    col.regions = pal_terrain_col,
    layer.name = "NDVI, 500 m buffer, parking",
    zcol = "ndvi_mean_wt")

#### The parking only----------
mv_union_sta_den_bg_int_prkng_tx_ndvi= den_bg_int_prkng_tx_ndvi %>% 
  st_transform(2876) %>% 
  st_simplify(dTolerance = 20) %>%  #make size much smaller for visual. intentionally ordered here
  st_intersection(union_station_2_mi_2876) %>% 
  st_make_valid() %>% 
  st_buffer(0) %>% 
  mapview(
    col.regions = pal_terrain_col,
    layer.name = "NDVI, parking only",
    zcol = "ndvi_mean_wt")

#### Visualize all at once--------------
mv_prkng_both = mv_union_sta_den_bg_int_prkng_comp_ndvi+ 
  mv_union_sta_den_bg_int_prkng_tx_ndvi

mv_prkng_both@map %>% 
  addFullscreenControl()
```

Due to space constraints, we cannot map each parking lot interactively on this website. Below is a static map of the parking lots in the Denver study area.

```{r parking-union,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
load("den_bg_int_prkng_tx_union.RData")
load("den_bg_int_prkng_comp_union.RData")
den_bg_int_prkng_tx_union %>% 
  ggplot()+
  geom_sf()+
  theme_bw()
```

# Supplementary tables
These tables are expanded versions of the corresponding main-text tables. They include a few more sub-scenarios, varying, for example, the percentage of land that would be converted to native plants (Scenarios 1 and 4) or the buffer size around the riparian areas (Scenario 2).

## Table 1: Description of scenarios (area, pop affecteds)
```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
load("hia_all_over_ndvi_over_equity_est_boot.RData")
load("hia_all_over_ndvi_by_equity_est_boot.RData")
load("lookup_scenario_sort_order.RData") #to get it to sort how I want
load("lookup_scenario_main_text.RData") #to restrict to those in the main text if necessary
load("lookup_scenario_num.RData")
load("lookup_scenario_label.RData")
hia_all_over_ndvi_over_equity_est_boot %>% 
  bind_rows(hia_all_over_ndvi_by_equity_est_boot) %>% 
  arrange(scenario_sort_order) %>% 
  left_join(lookup_scenario_main_text, by = c("scenario", "scenario_sub")) %>% 
#  filter(scenario_main_text==1) %>% #remove from appendix but include in main-text tables
  left_join(lookup_scenario_num , by = c("scenario", "scenario_sub")) %>% #add this for presentation
  left_join(lookup_scenario_label, by=  c("scenario", "scenario_sub")) %>% #add better scenario labels
  dplyr::select(
    scenario_num,
    scenario_label, 
    scenario_sub, #not presenting sub-scenarios in the main text table but yes in appendix
    starts_with("equity_nbhd_denver_tertile"),
    starts_with("area_mi2_bg_int_tx"),
    starts_with("area_mi2_bg_int_res"),
#    starts_with("ndvi_quo"),
#    starts_with("ndvi_mean_alt"),
    starts_with("pop_affected")
  ) %>%
  ungroup() %>% 
  knitr::kable(
    caption = "The cumulative area of the treatment, its corresponding residential buffer, and the population affected",
    booktabs = TRUE,
    format.args = list(big.mark = ","),
    digits = c(0, 0, 0, 0, 1, 1, 1, 1, 1, 1,0, 0,0),
    col.names = c(
            "Scenario num.",
      "Scenario",
      "Sub-scenario",
      "Equity Tertile",
      "Area, treatment (mi^2^), Med.",
      "Area, treatment (mi^2^), LL",
      "Area, treatment (mi^2^), UL",
      "Area, residential buffer (mi^2^), Med.",
      "Area, residential buffer (mi^2^), LL",
      "Area, residential buffer (mi^2^), UL",
      "Pop. affected, est.",
      "Pop. affected, LL",
      "Pop. affected, UL"
      )  )
```
## Table 2: NDVI at baseline, alternative, and difference therein
```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
load("hia_all_over_ndvi_over_equity_est_boot.RData")
load("hia_all_over_ndvi_by_equity_est_boot.RData")
load("lookup_scenario_sort_order.RData") #to get it to sort how I want
load("lookup_scenario_main_text.RData") #to restrict to those in the main text if necessary
load("lookup_scenario_num.RData")
hia_all_over_ndvi_over_equity_est_boot %>% 
  bind_rows(hia_all_over_ndvi_by_equity_est_boot) %>% 
  arrange(scenario_sort_order) %>% 
  #filter to just the main sub-scenarios to present in the main text
  left_join(lookup_scenario_main_text, by = c("scenario", "scenario_sub")) %>% 
#  filter(scenario_main_text==1) %>% #remove from appendix but include in main-text tables
  left_join(lookup_scenario_num , by = c("scenario", "scenario_sub")) %>% #add this for presentation
  left_join(lookup_scenario_label, by=  c("scenario", "scenario_sub")) %>% #add better scenario labels
  dplyr::select(
    scenario_num,
    scenario_label, 
    scenario_sub,
    starts_with("equity_nbhd_denver_tertile"),
    starts_with("ndvi_mean_wt_tx"),
    starts_with("ndvi_quo"),
    starts_with("ndvi_mean_alt"),
    starts_with("ndvi_diff")
  ) %>% 
  knitr::kable(
    caption = "NDVI before, after, and difference",
    booktabs = TRUE,
    col.names = c(
      "Scenario num.",
      "Scenario", 
      "Sub-scenario",
      "Equity Tertile", 
      "Baseline NDVI, tx area, Med.",
      "Baseline NDVI, tx area, LL",
      "Baseline NDVI, tx area, UL",
      "Baseline NDVI, res. area, Med.",
      "Baseline NDVI, res. area, LL",
      "Baseline NDVI, res. area, UL",
      "Alternative NDVI, res. area, Med.",
      "Alternative NDVI, res. area, LL",
      "Alternative NDVI, res. area, UL",
      "Difference: NDVI, res. area, Med.",
      "Difference: NDVI, res. area, LL",
      "Difference: NDVI, res. area, UL"
      ),
    format.args = list(big.mark = ","),
    digits = c(0,0,0,0, 2,2,2,2,2,2,2,2,2,3,3,3)) #max digits after decimal
```

## Table 3: Estimated deaths averted (total and rate per 100k) under each scenario
```{r,  eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
load("hia_all_over_ndvi_over_equity_est_boot.RData")
load("hia_all_over_ndvi_by_equity_est_boot.RData")
load("lookup_scenario_sort_order.RData") #to get it to sort how I want
load("lookup_scenario_main_text.RData") #to restrict to those in the main text if necessary
load("lookup_scenario_num.RData")
hia_all_over_ndvi_over_equity_est_boot %>% 
  bind_rows(hia_all_over_ndvi_by_equity_est_boot) %>% 
  arrange(scenario_sort_order) %>% 
#  filter(scenario_main_text==1) %>% #remove from appendix but include in main-text tables
  left_join(lookup_scenario_main_text, by = c("scenario", "scenario_sub")) %>% 
  left_join(lookup_scenario_num , by = c("scenario", "scenario_sub")) %>% #add this for presentation
  left_join(lookup_scenario_label, by=  c("scenario", "scenario_sub")) %>% #add better scenario labels
  dplyr::select(
    scenario_num,
    scenario_label, 
    scenario_sub, #in supp. tables but not main text.
    starts_with("equity_nbhd_denver_tertile"),
    starts_with("deaths_prevented_coun"),
    starts_with("deaths_prevented_per_pop_100k")
  ) %>% 
  knitr::kable(
    caption = "Estimated deaths averted",
    booktabs = TRUE,
    col.names = c(
      "Scenario num.",
      "Scenario", 
      "Sub-scenario",
      "Equity Tertile", 
      "Deaths prevented, est.", 
      "Deaths prevented, LL", 
      "Deaths prevented, UL",
      "Deaths prevented per 100k, est.", 
      "Deaths prevented per 100k, LL", 
      "Deaths prevented per 100k, UL"
      ),
    format.args = list(big.mark = ","),
    digits = c(0,0,0,0,0,0,0,0,0)) #max digits after decimal
```


# References
