---
title: "Test-analysis"
author: "Michael D Garber"
date: "March 21 2022"
---

I would like this to show up on the website, but I don't want it to be included in the nav page. I want the ability to link to it, though.

Test full-screen mapview per
https://github.com/r-spatial/mapview/issues/38

```{r loadpackages, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}
library(here)
library(tidyverse)
library(sf)
library(terra)
library(raster) 
library(mapview)
library(leaflet.extras)
```

# NDVI for the Denver area, 2016-2021
We obtained the normalized difference vegetation index (NDVI) from the Landsat-8 satellite at a resolution of 30 meters squared at a roughly 15-day interval between 2016 and 2021 for the following area around Denver:

## Define study area
```{r full-screen-test, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}
setwd(here("data-processed"))
load("den_co_tract_geo.RData") 
mv_den_co_tract_geo = den_co_tract_geo %>% 
  mapview(zcol = "tract_fips", layer.name = "Census Tract") 

mv_den_co_tract_geo@map %>% 
  addFullscreenControl()
```


